---
title: "Induction data"
author: "Charles Kemp"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, progress = TRUE)
```

```{r packageload, message=FALSE}
library(here)
library(tidyverse)
library(dplyr)
library(R.matlab)
library(janitor)
```


1) Data from Kemp, Charles, and Joshua B. Tenenbaum (2009) "Structured statistical models of inductive reasoning." Psychological Review 116(1): 20 - 58

```{r islandweb, message=FALSE}
d_mat <- readMat(here("data_old", "islandweb.mat"))

nodes <- array(unlist(d_mat$names))
nodes[7] <- "sandshark"

args <- as_tibble(d_mat$args, .name_repair = "universal") %>%  
  setNames(c("premise", "conclusion")) %>% 
  mutate(across(premise:conclusion, ~recode(., "1"=nodes[1], "2"=nodes[2], "3"=nodes[3], "4"=nodes[4], "5"=nodes[5], "6"=nodes[6], "7"=nodes[7]))) 


d1 <- args %>% 
  mutate(human_rating = d_mat$human[,1], predicate = unlist(d_mat$proptypes[1]))
d2 <- args %>% 
  mutate(human_rating = d_mat$human[,2], predicate = unlist(d_mat$proptypes[2]))
d3 <- args %>% 
  mutate(human_rating = d_mat$human[,3], predicate = unlist(d_mat$proptypes[3]))

data_release <- d2 %>% 
  rbind(d3) %>% 
  mutate(predicate = recode(predicate, prop="gene", dis="disease"))
  

write_csv(data_release, here("data_release", "data_islandfoodwebarguments.csv"))
```


```{r mammalsweb, message=FALSE}
d_mat <- readMat(here("data_old", "mammalsweb.mat"))

nodes <- array(unlist(d_mat$names))

args <- as_tibble(d_mat$args, .name_repair = "universal") %>%  
  setNames(c("premise", "conclusion")) %>% 
  mutate(across(premise:conclusion, ~recode(., "1"=nodes[1], "2"=nodes[2], "3"=nodes[3], "4"=nodes[4], "5"=nodes[5], "6"=nodes[6], "7"=nodes[7]))) 


d1 <- args %>% 
  mutate(human_rating = d_mat$human[,1], predicate = unlist(d_mat$proptypes[1]))
d2 <- args %>% 
  mutate(human_rating = d_mat$human[,2], predicate = unlist(d_mat$proptypes[2]))
d3 <- args %>% 
  mutate(human_rating = d_mat$human[,3], predicate = unlist(d_mat$proptypes[3]))

data_release <- d2 %>% 
  rbind(d3) %>% 
  mutate(predicate = recode(predicate, prop="gene", dis="disease"))

write_csv(data_release, here("data_release", "data_mammalsfoodwebarguments.csv"))

```

Data from  Kemp, C. & Tenenbaum, J. B. (2003). Theory-based induction. Proceedings of the 25th Annual Conference of the Cognitive Science Society.

```{r sanjana, message=FALSE}
d_mat <- readMat(here("data_old", "Sanjana.mat"))

nodes= c("horse", "cow", "chimp", "gorilla", "mouse", "squirrel", "dolphin", "seal", "elephant", "rhino") 
args <- as_tibble(d_mat$args, .name_repair = "universal") %>%  
  setNames(c("premise_1", "premise_2", "premise_3", "conclusion")) %>% 
  mutate(across(premise_1:conclusion, ~recode(., "0" = "NA", "1"=nodes[1], "2"=nodes[2], "3"=nodes[3], "4"=nodes[4], "5"=nodes[5], "6"=nodes[6], "7"=nodes[7], "8"=nodes[8], "9"=nodes[9], "10"=nodes[10]))) 

data_release <- args %>% 
  mutate(human_rating = d_mat$human)

write_csv(data_release, here("data_release", "data_sanjana.csv"))
```

Data from  Kemp, Charles. "Inductive reasoning about chimeric creatures." Advances in Neural Information Processing Systems 24 (2011): 316-324.

```{r chimeras_individual, message=FALSE}

stim <- readMat(here("data_old", "chimeraexpt_adjadjpairs.mat"))

features <- unlist(stim$features)
objects <- unlist(stim$objects)
args <-  as_tibble(stim$chosenpairs, .name_repair = "universal")
colnames(args) = c("premise_1", "premise_2")
singlemap <- as_tibble(stim$singlemap, .name_repair = "universal")
colnames(singlemap) = c("sm_premise_1", "sm_premise_2")
singlemap <- singlemap %>% 
  rowid_to_column(var="prempairid") 

d_chimera <- read_csv(here("data_old", "chimera_exptresults.csv")) %>% 
  clean_names() %>% 
  select(-sum_model_max_1) %>% 
  rename(human_rating = human_max_7, conclusion = test_feature, prem1_tmp = premise_1, prem2_tmp = premise_2 ) %>% 
  rowid_to_column(var="argument")  %>% 
  mutate(prempairid = (argument-1) %/% 10 + 1) %>% 
  left_join(singlemap, by = c("prempairid")) %>% 
  # allow for premise permutation in singlemap
  mutate(premise_1 =  case_when(sm_premise_1 == 1 ~ prem1_tmp, 
                                TRUE ~ prem2_tmp)) %>% 
  mutate(premise_2 =  case_when(sm_premise_1 == 1 ~ prem2_tmp, 
                                TRUE ~ prem1_tmp)) %>% 
  select(argument, premise_1, premise_2,  conclusion, edge_qn, conflict_qn)
  
read_chimera_individual <- function(filename, participantnum) {
  fullp <- readMat(here("data_old", "chimeras_individual", filename))
  p <- fullp$userps
  
  condnumber_index <- which(rownames(p) == "condnumber")
  gens_index <- which(rownames(p) == "gens")
  condnumber <- unlist(p[condnumber_index])
  gens<- matrix(unlist(p[gens_index]), 10) 
  
  gens <-t(gens[,2:41]) %>% 
    as_tibble() %>% 
    pivot_longer(V1:V10, names_to = "conclusion", values_to = "rating") %>% 
    rowid_to_column(var="argument") %>% 
    mutate(condition = condnumber, participant=participantnum)
}

files <- dir(path = here("data_old", "chimeras_individual"), "*.mat") %>% 
  as_tibble() %>% 
  rename(filename = value) %>% 
  rowid_to_column(var="participant_id")   
  
all_participants <- map2_df(files$filename, files$participant_id, read_chimera_individual)

all_participants_release <- all_participants %>% 
  select(-conclusion) %>% 
  right_join(d_chimera, by = "argument") %>% 
  mutate(condition = recode(condition, "1" = "both_premises", "2" = "premise_1", "3" = "premise_2"))
  
mean_data <- all_participants_release %>% 
  group_by(argument, premise_1, premise_2, conclusion, condition) %>% 
  summarize(human_rating = mean(rating))

write_csv(all_participants_release, here("data_release", "data_chimeras.csv"))
```


GRAVEYARD: made a plot to compare the average results for the both_premise condition with the average results in `here("data_old", "chimera_exptresults.csv"))`. The two sets of results are fairly close but not identical -- not sure why. Perhaps the averages in `chimera_exptresults.csv` were based on a near-final (but not final) version of the data set.

```{r chimeras, message=FALSE, eval = FALSE}

stim <- readMat(here("data_old", "chimeraexpt_adjadjpairs.mat"))

features <- unlist(stim$features)
objects <- unlist(stim$objects)
args <-  as_tibble(stim$chosenpairs, .name_repair = "universal")
colnames(args) = c("premise_1", "premise_2")
singlemap <- as_tibble(stim$singlemap, .name_repair = "universal")
colnames(singlemap) = c("sm_premise_1", "sm_premise_2")
singlemap <- singlemap %>% 
  rowid_to_column(var="prempairid") 

d_chimera <- read_csv(here("data_old", "chimera_exptresults.csv")) %>% 
  clean_names() %>% 
  select(-sum_model_max_1) %>% 
  rename(human_rating = human_max_7, conclusion = test_feature ) %>% 
  rowid_to_column(var="argument") %>% 
  mutate(prempairid = argument  %/% 10 + 1) %>% 
  left_join(singlemap, by = c("prempairid")) %>% 
  mutate()

dict_old <- 1:57
dict_new <- features
dict <- setNames(dict_new, dict_old)
args <- args %>% 
  mutate(premise_1 = recode(premise_1, !!!dict), premise_2 = recode(premise_2, !!!dict))

d_mat <- readMat(here("data_old", "chimeras_exptdata.mat"))
  
conclusions= c("c1", "c2", "c3", "c4", "c5", "c6", "c7", "c8", "c9", "c10")

d1 <- as_tibble(matrix(unlist(d_mat$humavg[[1]]), 40), ..name_repair = "universal") %>% 
  rowid_to_column(var="argument") %>% 
  pivot_longer(V1:V10, names_to = "conclusion", values_to = "hh") %>% 
  mutate(condition = "both_premises")

d2 <- as_tibble(matrix(unlist(d_mat$humavg[[2]]), 40), ..name_repair = "universal") %>% 
  rowid_to_column(var="argument") %>% 
  pivot_longer(V1:V10, names_to = "conclusion", values_to = "hh") %>% 
  mutate(condition = "premise_1")

d3 <- as_tibble(matrix(unlist(d_mat$humavg[[3]]), 40), ..name_repair = "universal") %>% 
  rowid_to_column(var="argument") %>% 
  pivot_longer(V1:V10, names_to = "conclusion", values_to = "hh") %>% 
  mutate(condition = "premise_2")
  
#  unlist("...1") 
#d_c2 <- d_mat$humavg[2]
#d_c3 <- d_mat$humavg[3]

bt <- bind_cols(d_chimera, d1)

comparevals <- bt %>% 
  ggplot(aes(x=human_rating, y = hh)) +
  geom_point()

comparevals  
```




